---
title: "Fitting A Linear Model"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Setup

```{r}
library(tidyverse)
library(patchwork) 
library(ggplot2)

# Change this to your week 4 folder
#setwd(r"(C:/Users/Valentin/Documents/GIT_REPS/TUHH/Applied-Statistics/week04)")
setwd(r"(C:\AsAd\Python_Related\Applied_Statistics\AppliedStatistics\week04)")
```

Note that when saving images, do it with this ratio to fit perfectly
into the TUHH template:

width = 8.74, height = 5.11

## Import Data

```{r}
video_game_data <- read.csv("Video_Games_Sales_as_at_22_Dec_2016.csv")

clean_data <- video_game_data |>
  select(User_Score, Critic_Score, Year_of_Release, Global_Sales) |>
  mutate(User_Score = as.numeric(User_Score)*10) |>
  mutate(Year_of_Release = as.numeric(Year_of_Release)) |>
  select(where(is.numeric)) |>
  drop_na()

# Row index of the game with max sales
max_index <- which.max(clean_data$Global_Sales)

# View the entire row
clean_data[max_index, ]


print(length(clean_data$User_Score))
```

## Data Visualization

filtering N/A:

\begin{align*}
N = 16719 \\
N_\text{filtered} = 6894
\end{align*}

### User and Critic Score

```{r}

long <- clean_data %>%
  select(Year_of_Release, Global_Sales, User_Score, Critic_Score) %>%
  pivot_longer(cols = c(User_Score, Critic_Score),
               names_to = "Type",
               values_to = "Score")

distribution_plot <- ggplot(long, aes(x = Score, fill = Type)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 100)) +
  labs(
    x = "Score",
    y = "Density",
    title = "Distribution of User vs Critic Scores"
  ) +
  scale_fill_manual(values = c("User_Score" = "skyblue", "Critic_Score" = "orange")) +
  theme_minimal()

ggsave("plots/distribution_plot.png", distribution_plot, width = 8.74, height = 5.11, dpi = 300)

distribution_plot


par(mfrow = c(1, 2))

qqnorm(clean_data$User_Score, main = "QQ Plot: User Score", ylim = c(0, 100))
qqline(clean_data$User_Score, col = "red")

qqnorm(clean_data$Critic_Score, main = "QQ Plot: Critic Score", ylim = c(0, 100))
qqline(clean_data$Critic_Score, col = "red")

width_px  <- 8.74 * 300
height_px <- 5.11 * 300  

dev.copy(png, "plots/QQplots_User_Critic.png", width = width_px, height = height_px, res = 300)
dev.off()
```

## Year of Release and Global Sales

```{r}

sales_release_plot <- ggplot(clean_data, aes(x = Year_of_Release, y = Global_Sales)) +
  geom_point(alpha = 0.3) +
  labs(x = "Year of Release",
       y = "Global Sales (millions)",
       title = "Global Sales vs Year of Release") +
  theme_minimal()


ggsave("plots/sales_release_plot.png", sales_release_plot, width = 8.74, height = 5.11, dpi = 300)

sales_release_plot
```

Incomplete List of Games, Most Games are not on there, such as the Pc
version of Minecraft missing and the list only including games up to
2016 The 2006 Outlier is Wii U Sports

## Rating per Year and Sales

```{r}

# Compute average per year
avg_year <- long %>%
  group_by(Year_of_Release, Type) %>%
  summarise(mean_score = mean(Score, na.rm = TRUE))

# Filter only User_Score
long_user <- long %>% filter(Type == "User_Score")
avg_year_user <- avg_year %>% filter(Type == "User_Score")

# Plot
score_over_year_plot <- ggplot(avg_year_user, aes(x = Year_of_Release, y = mean_score)) +
  # Background scatter
  geom_point(data = long_user, aes(x = Year_of_Release, y = Score),
             alpha = 0.2, size = 1, color = "grey") +
  # Average line
  geom_line(size = 1.2, color = "skyblue") +
  geom_point(size = 2, color = "skyblue") +
  labs(x = "Year of Release",
       y = "Average User Score",
       title = "Average User Scores Over Years") +
  theme_minimal()


ggsave("plots/score_over_year_plot.png", score_over_year_plot, width = 8.74, height = 5.11, dpi = 300)

score_over_year_plot

```

## Model

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \varepsilon
$$
$$
\begin{align*}
where\\
Y & : \text{User Score} \\
X_1 & : \text{Critic Score} \\
X_2 & : \text{Year of Release} \\
X_3 & : \text{Global Sales} \\
\varepsilon & : \text{Random error term}
\end{align*}
$$

```{r}
model <- lm(User_Score ~ ., data = clean_data)

summary(model)
```


```{r}
#Getting model coefficients and P values

round(summary(model)$coefficients[, c("Estimate", "Pr(>|t|)")],4)
```
$$
\[
\begin{array}{l r r}
\text{Coefficient} & \text{Estimate} & \text{p-value} \\
\hline
\text{Intercept} & 1699.0412 & 0 \\
\text{Critic\_Score} & 0.6117 & 0 \\
\text{Year\_of\_Release} & -0.8318 & 0 \\
\text{Global\_Sales} & -0.3728 & 0 \\
\end{array}
\]
$$


```{r}
# Dont know how relevant this is
plot(model)
```

## ChatGPT Analysis

This is the summary of a multiple linear regression where you model
User_Score using all other variables in clean_data:

User_Score ∼ Critic_Score + Year_of_Release + Global_Sales
User_Score∼Critic_Score+Year_of_Release+Global_Sales

I’ll break it down piece by piece and then interpret it in plain
language.

1️⃣ Model fit (big picture) R-squared = 0.3985

About 40% of the variation in User_Score is explained by the three
predictors.

This is moderate explanatory power for real-world data.

Adjusted R-squared = 0.3982

Very close to R² → adding predictors is justified.

F-statistic (p \< 2.2e-16)

The model as a whole is highly statistically significant

At least one predictor meaningfully explains User_Score

2️⃣ Residuals (errors) Min -6.76 1Q -0.54 Median 0.14 3Q 0.72 Max 3.89

Errors are centered close to 0 → good

Slight right skew (larger positive max)

Typical prediction error ≈ ±1.1 User_Score points

3️⃣ Coefficients (core interpretation) 🔹 Intercept: 169.9 (Intercept) =
169.9

The predicted User_Score when:

Critic_Score = 0

Year_of_Release = 0

Global_Sales = 0

Not meaningful in practice, but necessary for the model.

🔹 Critic_Score: +0.061 Estimate = 0.061 p \< 2e-16

📈 Interpretation:

For every 1-point increase in Critic_Score, User_Score increases by
\~0.06 points, holding other variables constant.

Very strong and significant predictor.

🔹 Year_of_Release: −0.083 Estimate = -0.083 p \< 2e-16

📉 Interpretation:

Each additional year is associated with a 0.083 decrease in User_Score.

Suggests older games have slightly higher user scores (possible
nostalgia or selection bias).

Statistically very strong.

🔹 Global_Sales: −0.037 Estimate = -0.037 p = 1.46e-07

📉 Interpretation:

For each additional million units sold, User_Score drops by \~0.037
points.

Small effect, but statistically significant.

Could reflect popular games appealing to broader (not more critical)
audiences.

4️⃣ Statistical significance

All predictors:

Have very low p-values

Are statistically significant at any conventional level

But ⚠️:

Statistical significance ≠ practical importance

Critic_Score matters much more than Global_Sales.

5️⃣ What this model says in plain English

User ratings are strongly aligned with critic ratings. Older games tend
to receive slightly higher user scores. Higher sales are weakly
associated with lower user scores. The model explains about 40% of user
score variability.

6️⃣ Important caveats

Correlation ≠ causation

Year_of_Release likely captures:

changes in rating culture

survivorship bias

Global_Sales is often right-skewed → log-transform may help

clean_data$log_sales <- log1p(clean_data$Global_Sales)

## The following code and text helps cover the part of the assignment that asks to assess the Nature of randomness underlying the error terms:

```{r}
# Extract residuals and fitted values
residuals <- resid(model)
fitted_values <- fitted(model)
mean(residuals)
```
What to put in the Presentation slides:
Title:Mean of Residuals

Mean of residuals: 2.63 × 10⁻¹⁵ (≈ 0)

Numerical zero due to floating-point precision

Confirms 𝐸(𝜀)=0

Model does not systematically over- or under-predict

Supports unbiased error assumption

What to say for this slide:
The mean residual is essentially zero, which confirms that the unexplained variation in the model is centered around zero. This means the model is unbiased on average.

Since the residuals are centered around zero on average, the next step is to check whether they behave randomly across the range of fitted values and whether their variance remains constant.



```{r}
residuals_vs_fitted_values_plot <- ggplot(data.frame(fitted_values, residuals),
       aes(x = fitted_values, y = residuals)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs Fitted Values"
  ) +
  theme_minimal()
ggsave("plots/residuals_vs_fitted_values.png", residuals_vs_fitted_values_plot, width = 8.74, height = 5.11, dpi = 300)
```
What to put in the Presentation slide:
Title: Randomness & homoscedasticity

[Picture of the graph]

Residuals centered around zero

No clear nonlinear pattern

Variance decreases with fitted values

Funnel shape → heteroscedasticity

Linear trend reasonable; inference with caution

What to say for this slide:
This plot shows the residuals plotted against the fitted values. The residuals are centered around zero and show no clear nonlinear pattern, which supports the linearity assumption. However, the spread of the residuals decreases as the fitted values increase, forming a funnel shape. This indicates heteroscedasticity, meaning the error variance is not constant. Despite this, the errors still appear random around zero, so the model remains useful for explanatory purposes.

So overall, the linear structure is appropriate, but the non-constant variance should be kept in mind when interpreting the results

```{r}
png("plots/qqplot_residuals.png", width = 8.74 * 300, height = 5.11 * 300, res = 300 )
qqnorm(residuals, main = "QQ Plot of Residuals")
qqline(residuals, col = "red")
dev.off()
```
What to put in the Presentation slide:
Title: Normality of the error terms

[Picture of the graph]

Assessment of normality of error terms

Residuals follow reference line closely

Mild deviations in the tails

Normality assumption approximately satisfied

Suitable for inference in large samples

What to say for this slide:
The Q–Q plot is used to assess whether the residuals follow a normal distribution. Here, the residuals lie close to the reference line, with only mild deviations in the tails. This suggests that the normality assumption for the error terms is approximately satisfied. Given the large sample size, these small deviations are not a major concern for inference.

Together with the mean residual and residuals vs fitted plots, this confirms that the main assumptions of the linear model are reasonably satisfied.
